Concurrent Programming with Harmony

### 2020-08-19 book/paper.tex
Now supports ->

Harmony does not have a \textbf{return} statement.  (You can assign a return value
to a method by setting the \textit{result} variable.)  Neither does it support
\textbf{break} or \textbf{continue} statements in loops.  One reason for their absence is
that, particularly in concurrent programming, such control flow directions are highly
error-prone.  It's easy to forget to, say, release a lock when returning a value in the
middle of a method, and a major source of bugs in practice.

@@

\autoref{fig:petersonmethods} again shows Peterson's algorithm,
but this time with methods defined to enter and exit the critical
section.
The name \textit{mutex} is often used to denote a variable or value
that is used for mutual exclusion.
\texttt{P\_mutex} is a method that returns a ``mutex,'' which, in this
case, is a dictionary that contains Peterson's Algorithm's state:
a turn variable and two flags.
Both methods \texttt{P\_enter} and \texttt{P\_exit} take two arguments:
a pointer to a mutex and the process identifier (0 or 1).
(\texttt{!}\textit{pm})\texttt{.turn} is the value of the \texttt{.turn} key
in the dictionary that \textit{pm} points to. 
Note that the parentheses in this expression
are needed, as \texttt{!}\textit{pm}\texttt{.turn} would wrongly evaluate
\texttt{!}(\textit{pm}\texttt{.turn}).

@@

\begin{figure}
\begin{code}
\harmonysource{lockintf}
\end{code}
\caption{The \texttt{Lock} interface and implementation in the \texttt{synch} module.}
\label{fig:spinlocks}
\end{figure}

@@

\begin{figure}
\begin{code}
\harmonysource{locksusp}
\end{code}
\caption{The \texttt{Lock} interface in the \texttt{synchS} module uses suspension.}
\label{fig:suspension}
\end{figure}

@@

\begin{figure}
\begin{code}
\harmonysource{xy}
\end{code}
\caption{Imcomplete code with desired invariant $x + y = 100$.}
\label{fig:xy}
\end{figure}

@@

\begin{figure}
\begin{code}
\harmonysource{queuetest}
\end{code}
\caption{Test program for a concurrent queue.}
\label{fig:queuetest}
\end{figure}

\begin{figure}
\begin{code}
\harmonysource{queue}
\end{code}
\caption{A basic concurrent queue data structure.}
\label{fig:queue}
\end{figure}

\begin{figure}
\begin{code}
\harmonysource{queueMS}
\end{code}
\caption{A queue with separate locks for enqueuing and dequeuing items.}
\label{fig:queueMS}
\end{figure}

@@

The \texttt{alloc} module
\index{alloc module}%
supports thread-safe (but not interrupt-safe) dynamic allocation of
memory locations.  There are just two methods:

@@

The usage is similar to \texttt{malloc} and \texttt{free} in C.
\texttt{malloc}() is specified to \texttt{None} when running out of
memory, although this is an impossible outcome in the current
implementation of the module.

>>>>

Harmony does not have a \textbf{return} statement.  (You can assign a return value
to a method by setting the \textit{result} variable.)  Neither does it support
\textbf{break} or \textbf{continue} statements in loops.  One reason for their absence is
that, particularly in concurrent programming, such control flow directions are highly
error-prone.  It's easy to forget to, say, release a lock when returning a value in the
middle of a method, a major source of bugs in practice.

@@

Often, pointers point to dictionaries, and so if $p$ is such a pointer,
then $!(p).\mathtt{field}$ would evaluate to the specified field in the dictionary.
Note that the parentheses in this expression
are needed, as \texttt{!}\textit{p}\texttt{.field} would wrongly evaluate
\texttt{!}(\textit{p}\texttt{.field}).
$!(p).\mathtt{field}$ is such a common expression that, like C, Harmony supports the
shorthand $p-$$>$\texttt{field}, which greatly improves readability.

\autoref{fig:petersonmethods} again shows Peterson's algorithm,
but this time with methods defined to enter and exit the critical
section.
The name \textit{mutex} is often used to denote a variable or value
that is used for mutual exclusion.
\texttt{P\_mutex} is a method that returns a ``mutex,'' which, in this
case, is a dictionary that contains Peterson's Algorithm's state:
a turn variable and two flags.
Both methods \texttt{P\_enter} and \texttt{P\_exit} take two arguments:
a pointer to a mutex and the process identifier (0 or 1).
$\mathit{pm}\rightarrow\mathtt{turn}$ is the value of the \texttt{.turn} key
in the dictionary that \textit{pm} points to.

@@

\begin{figure}
\begin{code}
\harmonysource{lockintf}
\end{code}
\caption{\harmonylink{modules/synch.hny} The \texttt{Lock} interface and implementation in the \texttt{synch} module.}
\label{fig:spinlocks}
\end{figure}

@@

\begin{figure}
\begin{code}
\harmonysource{locksusp}
\end{code}
\caption{\harmonylink{modules/syncS.hny} The \texttt{Lock} interface in the \texttt{synchS} module uses suspension.}
\label{fig:suspension}
\end{figure}

@@

\begin{figure}
\begin{code}
\harmonysource{xy}
\end{code}
\caption{\harmonylink{code/xy.hny} Incomplete code with desired invariant $x + y = 100$.}
\label{fig:xy}
\end{figure}

@@

\begin{figure}
\begin{code}
\harmonysource{queuetest}
\end{code}
\caption{\harmonylink{code/queuetest.hny} Test program for a concurrent queue.}
\label{fig:queuetest}
\end{figure}

\begin{figure}
\begin{code}
\harmonysource{queue}
\end{code}
\caption{\harmonylink{code/queue.hny}A basic concurrent queue data structure.}
\label{fig:queue}
\end{figure}

\begin{figure}
\begin{code}
\harmonysource{queueMS}
\end{code}
\caption{\harmonylink{code/queueMS.hny} A queue with separate locks for enqueuing and dequeuing items.}
\label{fig:queueMS}
\end{figure}

@@

The \texttt{alloc} module
\index{alloc module}%
supports thread-safe (but not interrupt-safe) dynamic allocation of
shared memory locations.  There are just two methods:

@@

The usage is similar to \texttt{malloc} and \texttt{free} in C.
\texttt{malloc}() is specified to return \texttt{None} when running out of
memory, although this is an impossible outcome in the current
implementation of the module.
